The Second Data Release of the INT Photometric H$\alpha$ Survey of the Northern Galactic Plane (IPHAS DR2)

### 2014-07-29 paper/referee/response-revision1.txt

[Referee]
> 4.b. Section 3.2 also specifies that an ellipticity limit of eps>0.3 was used.
> What is the justification of this limit?  For example, what is the potential impact on
> photometry and sensitivity at eps=0.25?

[Authors' response]
We have edited the third paragraph in Section 3.2 to explain that the choice
of our ellipticity limit at eps>0.3 is based on inspecting examples of 
high-ellipticity exposures, which revealed no evidence for degraded photometry
up to ellipticities less than 0.3. In fact high ellipticities are correlated 
with superior seeing, and therefore somewhat deeper frames.

[Referee]
> 5.a. Were the atmospheric extinction coefficients of 0.09 (r) and 0.05 (i) simply
> adopted, or was there any attempt to measure them?  Is there an estimate of
> how much the extinction may have varied with time, either during a night, or
> seasonally?

[Authors' response]
The atmospheric extinction coefficients adopted are standard default values 
for La Palma.  No direct measures or corrections were attempted, because our 
re-calibration procedure naturally absorbs extinction variations as part of the
APASS-based zeropoint estimation.

@@

[Referee]
> 5.b. The PERCORR factor in equation 1 is used to correct for "sporadic gain variation."
> What was the nature of the variations when non-zero?  How was it measured?
> What is the size of the correction when it was applied?

[Authors' response]
We have edited the text to clarify that the PERCORR factor is in fact
used to correct for the slight difference in internal gain computed using 
the twilight flats compared to the much redder typical astronomical object.  
It is computed by making a robust average of the dark sky levels measured on each
detector during an observing run.  In other words its a few% level update
on the internal gain values estimated from the flats.

### 2014-07-29 paper/iphas-dr2-rev1.tex
-

{ \bf
Source detection and parametrisation is then carried out
using the standard methods in use by CASU,
which can be summarised in four steps
\citep[a discussion on each of these steps
and related points can be found in][]{Irwin1985,Irwin1997}:
\begin{enumerate}
\item The local sky background is estimated
by computing the sigma-clipped median intensity
in a grid of 64$\times$64 pixel bins across the image,
which is robust against contaminating sources.
Bilinear interpolation is then used to obtain an estimate 
of the background level at each pixel.
\item Objects, or blends of objects, are detected
by identifying groups of 4 or more neighbouring pixels
in which the intensity exceeds the background level by $1.25\sigma$.
\item The object are parametrised:
positions are obtained by computing 
the intensity-weighted isophotal centre-of-gravity
of each object,
while photometry is obtained 
by measuring the intensity
in a series of soft-edged circular apertures of increasing diameter 
(1\farcs 2, 2\farcs 3, 3\farcs 3, 4\farcs 6 and 6\farcs 6).
\item Finally, the parametrisation of overlapping objects
is refined by simultaneous fitting of
soft-edged apertures to each blend,
which is essentially identical 
to multiple "profile" fitting.
\end{enumerate}
A discussion on each of these steps
can be found in \citep{Irwin1985,Irwin1997}.
We note that the parametrisation of single, unconfused objects
is naturally more reliable than that of
objects which had to be deblended.
In \S\ref{sec:catalogue} we will explain
how blended sources are flagged in the catalogue.

%The sentence about deblending could be misleading since the objects are
%detected using our standard deblending algorithm while the aperture flux
%estimates for overlapping images are computed by simultaneous fitting of
%soft-edged apertures.  During this multiple "profile" fitting the sky
%level is based on the interpolated global detector-level fit

%This approach works very well 
%across the vast majority of the survey area.  Nevertheless,
%the Galactic Plane contains crowded regions 
%with large numbers of overlapping sources
%or rapidly spatially-varying nebulosity,
%where aperture photometry will be compromised by frequent blending
%or poor background subtraction. 
}

Finally, an astrometric solution is determined
based on the Two-Micron All Sky Survey (2MASS) point source catalogue \citep{Skrutskie2006},
which itself is calibrated 
in the International Celestial Reference System (ICRS).
A provisional photometric calibration is also provided 
based on the average zeropoint
determined from sets of standard stars observed within the same night.
Sources are classified morphologically
-- stellar, extended or noise --
based on the curve-of-growth determined
from measuring the source intensity in a series of growing apertures.
Finally, the resulting source detection tables are stored 
in multi-extension FITS files.

At the time of preparing DR2,
the CASU pipeline had processed
74,195 single-band IPHAS exposures 
in which a total of 1.9~billion \emph{candidate} detections
were made at the default detection level of 1.25\,$\sigma$.
This total inevitably includes spurious objects, artefacts and
duplicate detections;
in \S\ref{sec:catalogue} we will explain
how these have been removed or flagged in the final catalogue.
{ \bf
The pipelined data set -- comprising 2.5~terabyte of FITS files --
was then transferred to the University of Hertfordshire
for the purpose of transforming the raw
detection tables into a proper source catalogue which
(i) is quality-controlled,
(ii) is homogeneously calibrated, and 
(iii) contains user-friendly columns and warning flags.
It is these post-processing steps which distinguish the present release
from the IDR, which enforced less stringent quality limits,
did not offer a global calibration,
and provided a catalogue in which duplicate detections
of unique sources were not flagged.
These new processing steps are explained next.}

>>>>

{ \bf
Object detection and parametrisation is then carried out
using the standard methods in use by CASU,
which can be summarised in four steps
\citep[a discussion on each of these steps
and related points can be found in][]{Irwin1985,Irwin1997}:
\begin{enumerate}
\item The local sky background is estimated
by computing the sigma-clipped median intensity
in a grid of 64$\times$64 pixel bins across the image.
Bilinear interpolation is then used to obtain an estimate 
of the background level at each pixel,
which is robust against contaminating sources.
\item Objects, or blends of objects, are detected
by identifying groups of 4 or more neighbouring pixels
in which the intensity exceeds the background level by at least $1.25\sigma$.
\item The object are parametrised:
positions are obtained by computing 
the intensity-weighted isophotal centre-of-gravity
of each object,
and photometry is obtained 
by measuring the intensity
in a series of soft-edged circular apertures of increasing diameter 
(1\farcs 2, 2\farcs 3, 3\farcs 3, 4\farcs 6 and 6\farcs 6).
Objects are also classified morphologically
-- stellar, extended or noise --
based on their intensity as a function of aperture size.
\item The parametrisation of overlapping objects
is refined by simultaneous fitting of
soft-edged apertures to each blend,
effectively carrying out profile fitting where necessary.
We note that the parametrisation of these blended objects is naturally
less reliable than that of single, unconfused sources,
but they are clearly flagged in the catalogue
(to be explained in \S\ref{sec:catalogue}).
\end{enumerate}

%This approach works very well 
%across the vast majority of the survey area.  Nevertheless,
%the Galactic Plane contains crowded regions 
%with large numbers of overlapping sources
%or rapidly spatially-varying nebulosity,
%where aperture photometry will be compromised by frequent blending
%or poor background subtraction.

Having carried out object detection and characterisation,
an astrometric solution is then fitted
based on the Two-Micron All Sky Survey (2MASS)
point source catalogue \citep{Skrutskie2006},
which itself is calibrated 
in the International Celestial Reference System (ICRS).
A provisional photometric calibration is also determined 
based on the average zeropoint
obtained from sets of standard stars observed within the same night.

At the time of preparing DR2,
the pipeline had processed
74,195 single-band IPHAS exposures 
in which a total of 1.9~billion \emph{candidate} detections
were made.
This total inevitably includes spurious objects, artefacts and
duplicate detections;
in \S\ref{sec:catalogue} we will explain
how these have been removed or flagged in the final catalogue.

The pipelined data set -- comprising 2.5~terabyte of FITS files --
was then transferred to the University of Hertfordshire
for the purpose of transforming the raw
detection tables into a proper source catalogue which
(i) is quality-controlled,
(ii) is homogeneously calibrated, and 
(iii) contains user-friendly columns and warning flags.
It is these post-processing steps which distinguish the present release
from the IDR, which enforced less stringent quality limits,
did not offer a global calibration,
and provided a catalogue in which duplicate detections
of unique sources were not flagged.
These new processing steps are explained next.}
